Two Site, Single Disturbance, MultiPool

### 2020-11-20 TwoSite_SingleDisturb_MultiPool.Rmd
two site model. Some text

## initial conditions
## spin up
ne = 500
met = sample.int(nd,ne,replace=TRUE)
par = sample.int(np,ne,replace=TRUE)
Na <- matrix(NA,ne,3)
for(i in 1:ne){
  Na[i,] = VSEM(as.numeric(Theta[par[i],]),PAR = PAR[met[i],])[nt,2:4]
}
Na.orig = Na
par(mfrow=c(1,3))
hist(Na[,1],main="Cleaf")
hist(Na[,2],main="Csoil")
hist(Na[,3],main="Cstem")
```

@@

```{r}
## hard coded 
Q = diag(c(0.1,0.1,0.1)) ## process error
R = diag(c(.5,10,1))      ## observation error
R = as.matrix(Matrix::bdiag(R,R))  ## nsite = 2
THETA <- as.matrix(Theta[par,]) ## ensemble parameters
THETA = abind::abind(THETA,THETA,along = 3)  ## n=2 sites
mu0 <- c(1,1)            ## mean leaf & stem biomass after disturbance
V0  <- diag(c(0.15,0.25))^2## var in post disturbance leaf and stem biomass
alloc.soil = 0.25        ## fraction of removed C that goes in the soil
p.dist = 0.05

@@

Na <- NaK <- abind::abind(Na.orig,
                          Na.orig[sample.int(ne,ne),],  ## resample IC for site 2
                          along = 3)
NfK = Nf = Na*NA
THETAK <- THETA
d.prior = matrix(nrow=ne,ncol=2)
muf <- matrix(NA,3*nsite,4)
pf  <- array(0,c(3*nsite,3*nsite,4)) ## 3rd variable is class
for(t in 1:NT){

@@

  plot(No[,i],
       #ylim = range(rbind(CIN,data$Y)),xlim=xlim,type='n',
       ylab=pool.names[i],xlab="time")
  ecoforecastR::ciEnvelope(1:ncol(CI.EnKF),CI.EnKF[1,],CI.EnKF[3,],col=ecoforecastR::col.alpha('#b2df8a',0.7))
  lines(CI.EnKF[2,],col="#b2df8a",lwd=2)
  ecoforecastR::ciEnvelope(1:ncol(CI.Na),CI.Na[1,],CI.Na[3,],col=ecoforecastR::col.alpha('blue',0.5))
  lines(CI.EnKF[2,],col="blue",lwd=2)
  abline(v=10)  
  lines(Ntrue[,i],lty=2,lwd=2)
}
```
### Ensemble adj at time of disturbance

>>>>

## initial conditions
## spin up
ne = 1000
met = sample.int(nd,ne,replace=TRUE)
par = sample.int(np,ne,replace=TRUE)
Na <- matrix(NA,ne,3)
for(i in 1:ne){
  Na[i,] = VSEM(as.numeric(Theta[par[i],]),PAR = PAR[met[i],])[nt,2:4]
}
Na.orig = Na
par(mfrow=c(1,3))
hist(Na[,1],main="Cleaf")
hist(Na[,2],main="Csoil")
hist(Na[,3],main="Cstem")
```

@@

```{r}
## hard coded 
Q = diag(c(0.1,0.1,0.1)) ## process error
R = diag(c(.5,10,1))      ## observation error
R = as.matrix(Matrix::bdiag(R,R))  ## nsite = 2
THETA <- as.matrix(Theta[par,]) ## ensemble parameters
THETA = abind::abind(THETA,THETA,along = 3)  ## n=2 sites
mu0 <- c(1,1)            ## mean leaf & stem biomass after disturbance
V0  <- diag(c(0.15,0.25))^2## var in post disturbance leaf and stem biomass
alloc.soil = 0.25        ## fraction of removed C that goes in the soil
p.dist = 0.025

@@

Na <- NaK <- abind::abind(Na.orig,
                          Na.orig[sample.int(ne,ne),],  ## resample IC for site 2
                          along = 3)
NfK = Nf = Na*NA
THETAK <- THETA
d.prior = matrix(nrow=ne,ncol=2)
muf <- matrix(NA,3*nsite,4)
pf  <- array(0,c(3*nsite,3*nsite,4)) ## 3rd variable is class
for(t in 1:10){

@@

  plot(No[,i],
       #ylim = range(rbind(CIN,data$Y)),xlim=xlim,type='n',
       ylab=pool.names[i],xlab="time")
  ecoforecastR::ciEnvelope(1:ncol(CI.EnKF),CI.EnKF[1,],CI.EnKF[3,],col=ecoforecastR::col.alpha('#b2df8a',0.7))
  lines(CI.EnKF[2,],col="#b2df8a",lwd=2)
  ecoforecastR::ciEnvelope(1:ncol(CI.Na),CI.Na[1,],CI.Na[3,],col=ecoforecastR::col.alpha('blue',0.5))
  lines(CI.Na[2,],col="blue",lwd=2)
  abline(v=10)  
  lines(Ntrue[,i],lty=2,lwd=2)
}
```
### Ensemble adj at time of disturbance

### 2020-11-20 methods.Rmd
two site model. Some text

3. Simulated data experiments (known disturbance)
  + x Single Site, Single Disturbance  (add a case where EnKF has disturb but it's more rare)
  + x SSSD Multi Pool
  + x SS Multi-Disturbance - distinct
  + x SS Multi-Disturbance w/ distinct, PFT switching
      * mention parallel PFT ensembles
      * need to protect against long-term PFT leakage; related to more general need to "look back"
  + Simple Multi-site: spatial cov
  + SSMD - initially overlapping but diverge
      * how to handle updating of assignment based on new information?
  + SSMD - transition probabilities change with state
  + rate of false pos/neg: affected by p.dist, R, Q, P0 -- how much do we need to simulate?
4. Real-world examples
  + Describe Model & PEcAn
  + Calibrate state-and-transition prior
    + Landtrendr
    ftp://islay.ceoas.oregonstate.edu/change_agent_attribution
  + Applications: PNW
    + Logging
    + Fire
    + Land use transition (forest -> non?)

>>>>

3. Simulated data experiments (known disturbance)
  + describe model
  + x Single Site, Single Disturbance  (add a case where EnKF has disturb but it's more rare)
  + x SSSD Multi Pool
  + x SS Multi-Disturbance - distinct
  + x SS Multi-Disturbance w/ distinct, PFT switching
      * mention parallel PFT ensembles
      * need to protect against long-term PFT leakage; related to more general need to "look back"
  + x Simple Multi-site: spatial cov
  + SSMD - initially overlapping but diverge
      * how to handle updating of assignment based on new information?
  + SSMD - transition probabilities change with state
  + rate of false pos/neg: affected by p.dist, R, Q, P0 -- how much do we need to simulate?
4. Real-world examples
  + Calibrate state-and-transition prior
    + Landtrendr
    http://emapr.ceoas.oregonstate.edu/pages/data/viz/index.html
    + Location:
    tile h03v04
xmin       : -2115585 
xmax       : -1965585 
ymin       : 2564805 
ymax       : 2714805 
Upper Left  (-2115585.000, 2714805.000) (123d 6'32.29"W, 44d41'17.09"N)
Lower Left  (-2115585.000, 2564805.000) (122d35' 7.51"W, 43d23'46.18"N)
Upper Right (-1965585.000, 2714805.000) (121d16'45.91"W, 45d 3'17.45"N)
Lower Right (-1965585.000, 2564805.000) (120d47'15.95"W, 43d45'18.18"N)
Center      (-2040585.000, 2639805.000) (121d56'25.42"W, 44d13'35.07"N)
> res(cover)
[1] 30 30
> crs(cover)
CRS arguments:
 +proj=aea +lat_0=23 +lon_0=-96 +lat_1=29.5 +lat_2=45.5 +x_0=0 +y_0=0 +datum=NAD83
+units=m +no_defs 
    + Biomass [@Kennedy2018]
      1990-2017
    + Cover
      1990-2017
    + Disturbance
      1984-2011
      - nearest neighbor resampled to same as cover
    + LAI??
  + Calibrate model
  + Applications: PNW
    + Logging
    + Fire
    + Land use transition (forest -> non?)
